Modeling browser-based distributed evolutionary computation systems

### 2015-02-08 modeling-volunteer-computing-gecco.tex
Getting closer to the end

The fact that it run correctly during a massive lightning talk means that this setup is a
valid, no-cost, platform for massive evolutionary algorithms. There
were several ways of introducing it: during a talk, telling the
audience to visit the web that was posted in the first slide, and
after the talk, inserting the URL in a tweet with a brief explanation
and several tweets more indicating what it was about and what it
did. If any question arose, it was answered as fast as possible. In
the first experiment, the median number of different IPs was 18, with
a maximum of 28 and a minimum of 6. In the second, the median was 12.5
with a maximum of 19 and minimum of 8. The histogram of different IPs
in each experiment is shown in figure \ref{fig:histo:ips}. Probably
the most remarkable thing is that a cluster of 6 computers can, as a
minimum, be gathered for a distributed computing experiment by just
announcing it as a tweet, but the fact that we can obtain more than a
dozen computers in more than half the cases is also remarkable, and
goes to prove the excess of computing resources that people,
willingly, lend to a simple experiment.

>>>>

The fact that it run correctly during a massive lightning talk means that this setup is a
valid, no-cost, platform for massive evolutionary algorithms. There
were several ways of introducing it: during a talk, telling the
audience to visit the web that was posted in the first slide, and
after the talk, inserting the URL in a tweet with a brief explanation
and several tweets more indicating what it was about and what it
did. If any question arose, it was answered as fast as possible. In
the first experiment, the median number of different IPs was 18, with
a maximum of 28 and a minimum of 6. In the second, the median was 12.5
with a maximum of 19 and minimum of 8. The histogram of different IPs
in each experiment is shown in figure \ref{fig:histo:ips}. Probably
the most remarkable thing is that a cluster of 6 computers can, as a
minimum, be gathered for a distributed computing experiment by just
announcing it as a tweet, but the fact that we can obtain more than a
dozen computers in more than half the cases is also remarkable, and
goes to prove the excess of computing resources that people,
willingly, lend to a simple experiment.
%
\begin{figure*}[htb]
        \centering
        \subfigure[First experiment, run 3]{
                \includegraphics[width=0.45\linewidth]{data/log-puts-1-exp7.png}
                \label{fig:d1}
        }
        \subfigure[Second experiment, run 7]{
                \includegraphics[width=0.45\linewidth]{data/log-puts-2-traps-eps-3.png}
                \label{fig:d2}
        }
        \caption{IPs and the number of {\tt PUT}s (one every $n$
          generations, 50 in the first experiment, 100 in the second
          experiment) they did, sorted. These correspond to the runs
          in each batch of experiments with the max number of
          different IPs. Please note the algorithmic scale in the $x$ axis.}\label{fig:dist}
\end{figure*}

Not all users contribute in the same way. The distribution for the
runs in each experiment with the most IPs is shown in Figure
\ref{fig:dist}, which chooses the experiment with the most number of
uses and ranks, tops to down, every IP with the number of PUTs (which
should be multiplied by the number of generations per PUT, 50 in the
first case, 100 in the second, although at this point it is not
important, since all nodes in the experiment do it after the same
number of generations) contributed to the experiment. We can observe
in both cases a power law, something already observed by \cite{agajaj}
and which implies that there is a fixed proportion between the number
of cycles (generations) contributed by the first and that contributed
by the second which is roughly the same as the one between the second
and third. Note also that, since the experiment is hosted in a PaaS,
the IPs listed are anonymous, with a 10 always as the first number.
%
\begin{figure*}[htb]
        \centering
        \includegraphics[width=0.8\linewidth]{data/time-vs-ips.png}
        \caption{Total experiment time (time of the last PUT - time of
        the first one) vs number of IPs participating in it.}\label{fig:t}
\end{figure*}
%
\begin{figure*}[htb]
        \centering
        \includegraphics[width=0.8\linewidth]{data/intervals-histo.png}
        \caption{Histogram of the time between two PUTs for all the
          experiments in the second batch. The $x$ axis is logarithmic
        and bin size is 0.5, which means that the first bin contains
        all intervals of less than 3 seconds.}\label{fig:intervals}
\end{figure*}

At any rate, this exponential trend in the first posts breaks down
after a number of IPs. That could be due to several possible factors:
persons checking the web for just an instant, late comers, or very
slow computers that are not able to lend more cycles to the
simulation. The first IP in both cases contributes on the order of a
hundred PUTs, that is, several thousands generations and several
hundred thousands evaluations, which, once again, is quite remarkable
for such a simple experiment announced by an ephemeral tweet. It
should be noted that, since the speed in the browser virtual machines
(Chrome, mainly, but also Firefox) is as fast as the node.js version,
a parallel version does not add overhead to the single-browser
version, which, as mentioned above, can run in an independent
way: a second browser always adds to the first, even if the total
number of evaluations and the time needed will not speed up in the
same way. However, even taking into account the stochastic nature of a
single evolutionary algorithm, what we are going to find is a high
dispersion of the times needed to reach a solution, which are plotted
in Figure \ref{fig:t} against the number of IPs. There is a wide
range of durations, although there is no clear decrease of the time
needed as the number of IPs used increase. However, we will have to
investigate further the speed ups achieved and if these are
algorithmic or due to the particular asynchronous and web-based
implementation. Figure \ref{fig:intervals} might help understand the
kind of raw material we are using: It plots an histogram of the time
needed by all experiments to process 100 generations, 25600
evaluations and the associated evolutionary geneticry. The vast
majority of browsers processes it in less than 3 seconds (logarithm(
time ) < 0.5, which is equivalent to 3.16). However, it is interesting
to note that there are some intervals that go up to the thousands of
seconds, for which we cannot really offer an explanation, other than
the user leaving the page and returning after a while. However, 75\%
of the cases take less than 4 seconds, which is thus a statistical
measure of the kind of performance we can expect from clients for this
type of problem. We can also conclude that 75\%, or a majority, will
be in the fastest tier, although there will be a small amount of them
that will be quite slow and will probably be impossible to accomodate
in the distributed computing experiment.
